A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-17 graph_syntax.tex
merged changes

@@

	\begin{description}
	\item[connection edge:] Exchange of information between two nodes. These edges 
	can either be fixed (can not be removed from graph) or free (can be removed). 
	\item [implicit edge:] Implicit passing of information from a driver node to a 
	  variable node. A single variable node can have many incomming implicit edges. These edges are 
	  free and can be added or removed from the graph. 
	\end{description}

@@

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision}\end{equation}
\end{description} 
These violations will be managed in order to seek a working problem formulation.

@@

\subsection{Analysis Blocks and Connections}
\label{ss:analysis blocks and connections}
Analysis tools take in a number input variables and then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by a group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig.~\ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. If constructing a weighted graph, the weight calculation edge 
provides the opportunity to computational cost. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed sub-graphs within a graph.

\begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=4in]{images/analysis_block}
    \end{center}
    \vspace{-10pt}
\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
\label{f:analysis block}
\end{figure}

@@

MDAO problems require that information be passed between sets of analyses. When 
information from the output of one analysis block is passed, or connected, to the 
input of another analysis block, a new connection edge is added connecting the two 
variable nodes involved in the exchange. These connection edges are free, unlike the edges 
within an analysis block, they could be added or removed depending on the needs
of a given design problem. In other words, free connection edges are created by 
engineers linking the output of one tool to the input of another one. Note that 
it is allowed for a single output to have outgoing connection edges to multiple 
downstream inputs.

@@

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. So we define an expression block, as the collection of variable and model 
nodes related to a given objective or constraint function. Fig. \ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

@@

	It is important to note that the coupling cycles do not contain 
	any driver nodes. Solver, optimizer, or other iterative process is 
	invovled in the coupling definition--though one will be necessary for a valid 
  PSG. Additionally, a coupling cycle has no inherant start or end. It would be
  valid to pick any node in the cycle as a starting point, and proceed around the
  loop untill you get back to your starting point. For the Sellar Problem, picking 
  $D_1$ as the starting point would yeild a problem as given in 
	Eqn.~\ref{eqn:simple}, while picking $D_2$ would yield the problem as givne in 
  Eqn.~\ref{eqn:simple2}.

@@

\subsection{Fundamental Problem Graph}
	\label{ss:FPG}
	We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
	\begin{enumerate}
	\item[(1)] $\displaystyle{I_F \subset I}$ such that the remaining conditions hold
	\item[(2)] $\displaystyle{ \forall i \in I_F, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
	\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
	\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
	\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	\end{enumerate}
	The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
	Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
	Requirement (3) stipulates that only the global inputs that are being used should be included. 
	Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
	Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper indegree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

	The set $C_F$ is a set containing connection edges representing connections. Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$. While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.

\subsection{Obtaining the Fundamental Problem Graph}
	\label{ss:obtaining FPG}
	In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all. Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. Then the problem is reduced to deciding which connection edges to remove.

	%Then, mathematically, the FPG starts with $\mathcal A_1 = \{1,\ldots,m\}$, $I_{F,1} = I$, and $C_{F,1} = C_M$.
	Then, mathematically, the FPG starts with $C_{F,0} = C_M$, $E_{F,0} = E_M$, $V_{F,0} = V_M$.
	The nodes where $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ are considered first.
	\begin{enumerate}
	\item The first step is to detect holes and disconnect the connection edges following them. These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs. The set of variable nodes which are holes is created as
	\begin{equation}
	H = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
	\end{equation}
	which is the set of variable nodes with fewer incoming edges than are allowed by the lower indegree limit.
	Then the updated set of edges is created by removing the first set of connection edges following the hole:
	\begin{equation}
	C_{F,1} = C_{F,0} \setminus \{(x,y) \in E_M \st x=E_M^3(v) \txt{ for } v \in H\},
	\end{equation}
	%deg-v is now recaculated for the new C. 
	and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=3in]{images/analysis_block_hole}
		\end{center}
		\vspace{-20pt}
	\caption{Example variable node indicating a hole.}
	\label{f:hole}
	\end{figure}

	\item The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. The set of variable nodes which contain collisions is created as
	\begin{equation}
	S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
	\end{equation}
	For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
	\begin{equation}
	S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \st y \in S_\txt{nodes} \big \}
	\end{equation}
	Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
	An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
	We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
	Then we may construct sets of edges as
	\begin{equation}
	B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
	\end{equation}
	which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper indegree limit of $v_j$. The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
	\begin{equation}
	C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
	\end{equation}
	\begin{equation}
	E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
	\end{equation}
	and
	\begin{equation}
	F_2 = (V_M,E_{F,2})
	\end{equation}
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=3in]{images/analysis_block_collision}
		\end{center}
		\vspace{-20pt}
	\caption{Example variable node indicating a collision.}
	\label{f:collision}
	\end{figure}

	\item The third and final step is to remove any unused nodes and edges. The indexing set for only the analysis blocks with at least one used local output may be constructed as ((pruning))
	\begin{equation}
	I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
	\end{equation}
	where the degree is calculated with respect to $F_2$. This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.

	The set of used global inputs is constructed as
	\begin{equation}
	V_{\txt{in},F} = \Big\{v \in V_\txt{in} \ \Big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \Big\},
	\end{equation}
	which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.

	Finally, the set of nodes and edges describing a fundamental problem formulation is given by
	\begin{equation}
	V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
	\end{equation}
	\begin{equation}
	C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
	\end{equation}
	\begin{equation}
	E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
	\end{equation}
	\begin{equation}
	F = (V_F,E_F).
	\end{equation}

	\end{enumerate}

>>>>

	A \emph{path} $P=(V,E)$ from $x_0$ to $x_k$ in graph $G$ is a subgraph of $G$ with $V = \{x_0,x_1,\ldots,x_k\}$ and $E = \{(x_0,x_1),(x_1,x_2),\ldots,(x_{k-1},x_k)\}$.
	A \emph{reverse path} $P_R$ in graph $G$ is a path on $R$, where $R$ is the reverse graph of $G$ obtained by switching the orientation of every edge.

@@

	\begin{description}
	\item[connection edge:] Exchange of information between two nodes. These edges 
	can either be fixed (can not be removed from graph) or free (can be removed). ((this needs to be modified, as analysis blocks cannot be altered but they can be removed as a group))
	\item [implicit edge:] Implicit passing of information from a driver node to a 
	  variable node. A single variable node can have many incomming implicit edges. These edges are 
	  free and can be added or removed from the graph. 
	\end{description}

@@

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \label{e:collision}\end{equation}
\end{description} 
	An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.

@@

\subsection{Analysis Blocks and Connections}
\label{ss:analysis blocks and connections}
Analysis tools take in a number input variables and then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by a group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig.~\ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. If constructing a weighted graph, the weight calculation edge 
provides the opportunity to computational cost. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
	the edges within them are fixed, they become fixed sub-graphs within a graph. In this way, they cannot be altered but they can be removed as a group.
\begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=4in]{images/analysis_block}
    \end{center}
    \vspace{-10pt}
\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
\label{f:analysis block}
\end{figure}

@@

MDAO problems require that information be passed between sets of analyses. When 
information from the output of one analysis block is passed, or connected, to the 
input of another analysis block, a new connection edge is added connecting the two 
variable nodes involved in the exchange. These connection edges are free, unlike the edges 
within an analysis block, can be added or removed depending on the needs
of a given design problem. In other words, free connection edges are created by 
engineers linking the output of one tool to the input of another one. Note that 
it is allowed for a single output to have outgoing connection edges to multiple 
downstream inputs.

@@

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
	the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
	nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

@@

	It is important to note that the coupling cycles do not contain 
	any driver nodes. Solver, optimizer, or other iterative process is 
	invovled in the coupling definition--though one will be necessary for a valid 
  PSG. Additionally, a coupling cycle has no inherant start or end. It would be
  valid to pick any node in the cycle as a starting point, and proceed around the
  loop untill you get back to your starting point. For the Sellar Problem, picking 
  $D_1$ as the starting point would yeild a problem as given in 
	Eqn.~\ref{eqn:simple}, while picking $D_2$ would yield the problem as given in 
  Eqn.~\ref{eqn:simple2}.

@@

\subsection{Fundamental Problem Graph}
	\label{ss:FPG}
	We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
	\begin{enumerate}
	\item[(1)] $F \subset M$ and $V_\txt{out} \subset V_F$
	\item[(2)] $\displaystyle{\forall i \in I, \txt{ if } F \cap A_i \neq \emptyset \txt{ then } A_i \subset F}$
	\item[(3)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	\item[(4)] $\displaystyle{\forall v \in V_F \txt{ there exists a reverse path } P \subset R_F \txt{ from } x \txt{ to } v \txt{ with } x\in V_\txt{out}}$
	\end{enumerate}
	%The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
	%Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
	%Requirement (3) stipulates that only the global inputs that are being used should be included. 
	%Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
	Requirement (1) asserts that only the nodes and edges provided by the maximal connectivity graph can be used in the fundamental problem graph and that every global output must be included. 
	Requirement (2) requires that each analysis block must be entirely present or entirely absent. 
	Requirement (3) stipulates that the number of edges directed into each variable node must be within the lower and upper in-degree limits; 
	if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is a \emph{collision}.
	Lastly, requirement (4) ensures that only the nodes that are being used are included in the FPG by requiring that for every node a reverse path exists from at least one global output to that node. 
	The reverse graph $R_F$ is obtained from $F$ by simply switching the orientation of every edge in $E_F$.

	%The set $C_F$ is a set containing connection edges representing connections. Since $C_M$ is the set of all potential connections, we must have $C_F \subset C_M$. 
	%While no other conditions are explicitly stated for $C_F$, requirement (6) is actually a requirement on both $I_F$ and $C_F$.

\subsection{Obtaining the Fundamental Problem Graph}
	\label{ss:obtaining FPG}
	In general, there may be multiple different graphs that satisfy the FPG conditions in Sec.~\ref{ss:FPG}, though there may be none at all. 
	Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. 
	Then the problem is reduced to deciding which connection edges to remove.

((describe pruning as a subroutine enforcing step 4 and then step 2))

	\begin{description}
	\item[Subroutine: Pruning] 
		Given a digraph $G_1 = (V_1,E_1)$, let the subroutine denoted as $P_\txt{prune}$ operate on $G_1$ to produce a digraph $G_2 = (V_2,E_2)$ that satisfies requirements (2) 
		and (4), and write $G_2 = P_\txt{prune} ( G_1 )$. 
		This is accomplished by first creating the reverse graph of $G_1$, $R = (V_1,E_R)$. Then adding a new node $x$ to $V_R$ and adding edges directed from this node to each of the global outputs, i.e.,
	\begin{equation}
		\forall y \in V_\txt{out}, (x,y) \in E_R
	\end{equation}
		The set of nodes that may be reached from the global outputs is constructed as
	\begin{equation}
		U = \{ n \in V_1 \st \exists P \txt{ a path from $x$ to $n$}, \ P \subset R \}.
		\end{equation}
		The list of analysis blocks with at least one node that may be reach is constructed as
		\begin{equation}
		I_1 = \{ i \in I \st V_{A_i} \cap U \neq \emptyset \}
		\end{equation}
		The set of nodes to remove is
		\begin{equation}
		N = \{ v \in V_1 \st v \in V_{A_i} \txt{ for } i \in I_1, \txt{ or } v \in V_\txt{in}\setminus U \},
		\end{equation}
		which means that $N$ is composed of unused analysis block nodes and unused global input nodes. Then the new set of nodes is created as
		\begin{equation}
		V_2 = V_1 \setminus N,
		\end{equation}
		and edges involving the removed nodes are also deleted
		\begin{equation}
		E_2 = E_1 \setminus \{ (a,b) \st a \in N \txt{ or } b \in N  \}.
		\end{equation}
		The set of connection edges can be obtained as
		\begin{equation}
		C_2 = \{(x,y) \in V_{2} \st \sim(x \in V_{A_i} \txt{ and } y \in V_{A_i} \txt{ for some } i \in I_1)\},
		\end{equation}
		which means $C_2 \cap V_{A_i} = \emptyset  \ \forall i \in I_1$, and $E_2 = C_2 \cup \left( \bigcup_{i \in I_1} E_{A_i} \right)$.

	\item[Step 1: Holes] 
		The first step is to detect holes and disconnect the connection edges following them. 
		These connection edges are removed because they represent variables which cannot be determined because the analysis function does not have adequate inputs.

		To start, the an initial graph is created as a pruned copy of the MCG
		\begin{equation}
		F_0 = P_\txt{prune}(M),
		\end{equation}
		where $F_0 = (V_{F,0},E_{F,0})$, and $C_{F,0}$ is also obtained as described. 
		The set of variable nodes which are holes is created as
		\begin{equation}
		H = \{v \in V_{F,0} \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) < \txt{deg}_l^-(v) \},
		\end{equation}
		which is the set of variable nodes with fewer incoming edges than are allowed by the lower in-degree limit.
		Then the updated set of edges is created by removing the edges preceding or succeeding the analysis block:
		\begin{equation}
		C_{F,1} = C_{F,0} \setminus \{(x,y) \in C_{F,0} \st x \in V_{A_i} \txt{ or } y \in V_{A_i}, \txt{ and } H\cap V_{A_i} \neq \emptyset \},
	\end{equation}
	%deg-v is now recaculated for the new C. 
	and this step is demonstrated by Fig.~\ref{f:hole}. Because removing these edges can create new holes, this step must be repeated until no more holes are found.
		If this step every finds a global output to be a hole, meaning $V_\txt{out} \cap H \neq \emptyset$, then the algorithm stops because an FPG cannot be obtained. 
		Otherwise, it is guaranteed that an FPG can be obtained.
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=3in]{images/analysis_block_hole}
		\end{center}
		\vspace{-20pt}
	\caption{Example variable node indicating a hole.}
	\label{f:hole}
	\end{figure}

	\item[Step 2: Collisions] 
		The second step is to detect collisions and to then disconnect enough connection edges so that the collisions are resolved. 
		The set of variable nodes which contain collisions is created as
	\begin{equation}
	S_\txt{nodes} = \{v \in V_M \st t_\txt{node}(v) = \txt{`variable'} \txt{ and } \txt{deg}^-(v) > \txt{deg}_u^-(v) \}.
	\end{equation}
	For each collision node we can construct a set contained the edges directed in. The set containing all of these sets is constructed as
	\begin{equation}
		S_\txt{edges} = \big \{ \{(x,y) \in E_M\} \ \big| \ y \in S_\txt{nodes} \big \}
	\end{equation}
	Let $J=\{1,2,\ldots,|S_\txt{edges}|\}$ be an indexing set for $S_\txt{edges}$ such that each $S_{\txt{edges},j}$ corresponds to a set in $S_\txt{edges}$ for $j \in J$. 
	An example collision is shown in Fig.~\ref{f:collision} to indicate the definition of $S_{\txt{edges},j}$. 
	We can assume that $J$ also indexes $S_\txt{nodes}$ because there is a one--to--one correspondence between the elements in $S_\txt{nodes}$ and the elements in $S_\txt{edges}$ (which are sets). 
	Then we may construct sets of edges as
	\begin{equation}
	B_j = \big \{e_k \in S_{\txt{edges},j} \st k \in \{1,2,\ldots,K\} \txt{ with } K \leq \txt{deg}_u^-(v_j) \big \}, \ j \in J,
	\end{equation}
		which means that each set $B_j$ is constructed from the set $S_{\txt{edges},j}$ by taking only as many edges as are allowed by the upper in--degree limit of $v_j$. 
		The construction of each $B_j$ corresponds to making a decision on which edges to include and which edges not to include. Then let
	\begin{equation}
	C_{F,2} = \{ e \in C_{F,1} \st e \in B_j \txt{ for some } j \in J\},
	\end{equation}
	\begin{equation}
	E_{F,2} = E_M \setminus (C_M \setminus C_{F,2})
	\end{equation}
	and
	\begin{equation}
	F_2 = (V_M,E_{F,2})
	\end{equation}
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=3in]{images/analysis_block_collision}
		\end{center}
		\vspace{-20pt}
	\caption{Example variable node indicating a collision.}
	\label{f:collision}
	\end{figure}

	\item[Step 3: Finalize] 
		The third and final step is to prune the graph one last time
	\begin{equation}
		F = P_\txt{prune}(F_2),
	\end{equation}
		and $F = (V_F,E_F)$.
		The indices of the used analysis blocks can then be found as
	\begin{equation}
		I_F = \{ i \in I \st A_i \cap F \neq \emptyset \}
	\end{equation}
%		\begin{
%		\begin{equation}
%		I_F =  \{i \in I \ | \ \exists v \in V_{A_i} \txt{ such that }  t_\txt{node}(E_{F,2}^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v) > 0 \},
%		\end{equation}
%		where the degree is calculated with respect to $F_2$. 
%		This set excludes both the analysis blocks that were unused in the original MCG and those that became unused in step 1 or step 2.
%
%		The set of used global inputs is constructed as
%		\begin{equation}
%		V_{\txt{in},F} = \big\{v \in V_\txt{in} \ \big| \ |\{(x,y) \in E_{F,2}(v) \st y \in V_{A_i} \txt{ for } i \in I_F  \}| > 1 \big\},
%		\end{equation}
%		which requires that at least two edges directed out of the nodes in $V_{\txt{in},F}$ must be to the analysis blocks indexed by $I_F$.
%
%		Finally, the set of nodes and edges describing a fundamental problem graph is given by
%		\begin{equation}
%		V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right),
%		\end{equation}
%		\begin{equation}
%		C_{F} = \{ (x,y) \in C_{F,2} \st x,y \in V_F\},
%		\end{equation}
%		\begin{equation}
%		E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right),
%		\end{equation}
%		\begin{equation}
%		F = (V_F,E_F).
%		\end{equation}

	\end{description}

This process will always provide an FPG if one exists. If an FPG does not exist, this process will reveal the limiting factors preventing a valid formulation.
